Finite-state morphological transducers for three Kypchak languages

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
some cleaning up and rearranging and adding refs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48177 72bbbca6-d526-0410-a7d9-f06f51895060

\setromanfont{Times New Roman}

@@

These languages belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.  This branch can be divided into three subbranches.  Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.  As such, each of these three languages represents a different one of the three branches of Kypchak.  The geographic distribution of the languages is shown in map \ref{map}.

\begin{map*}
	\label{map}
	\includegraphics[width=\textwidth]{map/map}
	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
\end{map*}

In a linguistic sense, these languages have different amounts of influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.

@@

\subsection{Kazakh}
Kazakh /qɑzɑq/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.  Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.  Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{encyclopedin} people.

\subsection{Tatar}
Tatar /tɑtɑr/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.  It is co-official with Russian in Tatarstan --- a republic within Russia.  A majority of native speakers of both languages are bilingual in Russian. %, and the standard varieties our system FIXMEs are written in Cyrillic.

>>>>

\defaultfontfeatures{Mapping=tex-text}
\setromanfont{Times New Roman}
\newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}

@@

@@

The three languages for which transducers were developed belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.  This branch can be divided into three subbranches.  Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.  As such, each of these three languages represents a different one of the three branches of Kypchak.  The geographic distribution of the languages is shown in map \ref{map1}.

\begin{map*}[htbp]
	\includegraphics[width=\textwidth]{map/map}
	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
	\label{map1}
\end{map*}

These languages have different amounts of linguistic influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.

\subsection{Kazakh}
Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.  Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.  Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{encyclopedin} people.

\subsection{Tatar}
Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.  It is co-official with Russian in Tatarstan --- a republic within Russia.  A majority of native speakers of both languages are bilingual in Russian. %, and the standard varieties our system FIXMEs are written in Cyrillic.
